\documentclass[t]{beamer}
\usepackage[utf8]{inputenc}

\input{../common/setup.tex}

\title{Scala reflection and macros}

\author{Mikhail Limanskiy}

\institute{SymphonyTeleca}

\date{\today}

\begin{document}

\begin{frame}
    \titlepage
\end{frame}

\section{Reflection}

\begin{frame}[fragile]
\frametitle{Java\texttrademark generics limitations}

In Java all generic type information is removed in compile time \cite{erasure}.  As result, there no type info for
parametrized types in JVM.  As result you cannot use type parameters for type checks or instantiation:

\begin{lstlisting}[language=Java]
public <E> void append(List<E> list, Class<E> cls) throws Exception {
//  E elem = new E(); // will not compile
    E elem = cls.newInstance();   // OK
    list.add(elem);
}

public <T extends Foo> T getByType(List<Foo> foos, Class<T> clazz) {
    List<T> result = new ArrayList<>();
    for (Foo foo : foos) {
//      if (foo instanceOf T) result.add(foo);  // will not compile
        if (clazz.instanceOf(foo)) result.add(foo);
    }
    return result;
}
\end{lstlisting}
\end{frame}

\begin{frame}[fragile]
\frametitle{Type erasure affects Scala}
This code will compile with warnings, but not work:
\begin{lstlisting}
def foo(item: Option[Any]) = item match {
  case t: Option[Int] => println("int")
  case t: Option[Boolean] => println("bool")
  case _ => println("other")
}
foo(Some(false))
\end{lstlisting}
\begin{lstlisting}[breaklines=true]
<console>:8: warning: non-variable type argument Int in type pattern Option[Int] is unchecked since it is eliminated by erasure
  case t: Option[Int] => println("int")
          ^
<console>:9: warning: non-variable type argument String in type pattern Option[String] is unchecked since it is eliminated by erasure
  case t: Option[String] => println("bool")
          ^
<console>:9: warning: unreachable code
  case t: Option[String] => println("bool")
                                   ^
\end{lstlisting}
\end{frame}

\begin{frame}[fragile]
\frametitle{Reflection universes}
Scala reflection API contains two universes: Runtime and Compile-time.

\begin{itemize}
\item Runtime reflection is used to get class information at program runtime. It's a same type of reflection as in Java.
\item Compile-time reflection allows you to get type information during compilation, and is used with Scala Macros.
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Obtaining runtime type}
\begin{lstlisting}
def foo[T: TypeTag](item: Option[T]) = {
  val tt = typeOf[T]
  item match {
    case t if tt =:= typeOf[Boolean] => println("a bool")
    case t if tt =:= typeOf[Int]     => println("an int")
    case t if tt <:< typeOf[Seq[_]]  => println(s"a sequence ($tt)")
    case _                           => println("other")
  }
}

foo(Some(5)) // prints "a bool"
foo(Some(List(5,2))) // prints "a sequence (List[Int])"
\end{lstlisting}
\end{frame}

\begin{frame}[fragile]
\frametitle{Macros introduction}
Scala macros provide ability to generate and transform code at compile time.  You can think what macro is a function which
takes \strong{Abstract Syntax Tree} (AST) and returns another AST.

Let's define a simple ``Hello world'' macro:

\begin{lstlisting}
import scala.language.experimental.macros
import scala.reflect.macros.blackbox.Context

object HelloWorld {
  def hello: Unit = macro hello_impl

  def hello_impl(c: Context): c.Tree = {
    import c.universe._
    Apply(
      Ident(
        TermName("println")
      ),
      List(
        Literal(
          Constant("Hello world")
        )
      )
    )
  }
}
\end{lstlisting}

Now we can use it as any other function:

\begin{lstlisting}
HelloWorld.hello
\end{lstlisting}

\end{frame}

\begin{frame}[fragile]
\frametitle{Reify and quasiquotes}
Fortunately usually you don't need to construct syntax trees by hand.

You can use function \linline{reify} to get an AST:
\begin{lstlisting}
def hello_impl(c: Context): c.Expr[Unit] = {
  import c.universe._

  reify {
    println("Hello world")
  }
}
\end{lstlisting}

Another option is to construct AST from string using quasiquotes:
\begin{lstlisting}
def hello_impl(c: Context): c.Tree = {
  import c.universe._

  q"""
    println("Hello world")
  """
}
\end{lstlisting}
\end{frame}

\begin{frame}[fragile]
\frametitle{Writing jsonizer}
Assume we want to develop a toJson method for case classes. Something like:
\begin{lstlisting}
case class Person(name: String, age: Int) extends Jsonable

val p = Person("John Doe", 42)

println(p.toJson) // { "name" : "John Doe", "age" : 42 }
\end{lstlisting}

Let's start with defining required classes:
\begin{lstlisting}
trait Jsonable

object Jsonizer {
  implicit class Jsonized[T <: Jsonable](val data: T) extends AnyVal {
    def toJson: String = macro toJsonImpl[T]
  }

  def toJsonImpl[T: c.WeakTypeTag](c: Context): c.Expr[String] = {
    import c.universe._
    c.Expr[String](Literal(Constant("{}")))
  }
}
\end{lstlisting}

\end{frame}

\begin{frame}{Bibliography}
\begin{thebibliography}{00}
\bibitem{scala}Martin Odersky, Lex Spoon, Bill Venners:
\emph{Programming in Scala},
Artima, 2nd edition, 2011
\bibitem{erasure}Java\texttrademark  Tutorials: \emph{Type erasure}, \url{http://docs.oracle.com/javase/tutorial/java/generics/erasure.html}
\bibitem{reflect}Heather Miller, Eugene Burmako, Philipp Haller: \emph{Scala reflection tutorial}, \url{http://docs.scala-lang.org/overviews/reflection/overview.html}
\bibitem{macro}Eugene Burmako, \emph{Scala macros tutorial}, \url{http://docs.scala-lang.org/overviews/macros/usecases.html}
\bibitem{quasiquotes}Denys Shabalin, \emph{Quasiqoutes tutorial}, \url{http://docs.scala-lang.org/overviews/quasiquotes/intro.html}
\end{thebibliography}
\end{frame}

\input{../common/theend.tex}

\end{document}
